Say Cheese! A Snapshot of Canadian Family Demographics

### 2022-03-20 outputs/paper/paper.Rmd
updated graphs and some edits to results + discussion

---
title: "Say Cheese! A Snapshot of Canadian Family Demographics"
subtitle: "An analysis of insights from Canada's 2017 General Social Survey on Family"
author: 
  - Kimlin Chin
  - Jack McKay
  - Jacob Yoke Hong Si
thanks: "Code and data are available at: https://github.com/jmacattack27/General-Social-Survey-on-Family-2017- ."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "What does the average family look like in Canada? Using data from the 2017 Canadian General Social Survey (GSS) on Family, we look to paint a picture of Canadian families through the lens of family size and type, marriage and children, place of origin, income, religion and mental health. We found that most families are small, well-off and in good mental health, and that immigrants make up about one fifth of the population, with a propensity to settle in larger cities and urban centers. We also discuss possible factors that may have influenced this family image, including Canada's investment in family planning and immigration levels planning."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

```{r agesrespondents, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Respondents' Ages"}

@@

# Plot for total number of children
five_year_groupings %>%
  ggplot(aes(x = age_group_5)) +
  geom_bar() +
  # scale_y_continuous(breaks = seq(0, 7000, by = 1000)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 50, hjust=1)) +
  labs(title = "Distribution of respondents' ages",
       x = "Five-year Age groups",
       y = "Number of responses")

@@

@@

Figure \@ref(fig:numchildren) shows the distribution of the total number of children per respondent. Note that this is not *per household* i.e. persons living together, but rather is the number of children had by the respondent of the household throughout their lifetime. Here we see that the majority of respondents either had 0 children or 2, with almost an identical number of respondents with 0 children as there are with 2. Similarly, the number of respondents with 1 child is very close to the number of those with 3. Thus, we note that about `r round(zero_children$total_percent,2)`%of Canadians do not have any children, and those who do usually have 1-3 children.

```{r numchildren, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Total Number of Children per Respondent"}
# Plot for total number of children
child_data %>%
  select(total_children1) %>%
  ggplot(aes(x = total_children1)) +
  geom_bar(fill = "#377EB8") +
  scale_y_continuous(breaks = seq(0, 7000, by = 1000)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "Total Number of Children per Respondent",
       x = "Number of Children",
       y = "Number of Responses") +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r agemarbirth, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Ages at Marriage and Birth"}
# Plot for age at first marriage
marriage_plot<- gss_data %>%
  mutate(age_at_first_marriage = floor(age_at_first_marriage)) %>%
  select(age_at_first_marriage) %>%
  ggplot(aes(x = age_at_first_marriage)) +
  geom_bar(width = 0.95, fill = "#E41A1C") +
  scale_x_continuous(limits = c(15, 50), breaks = c(15, seq(0, 50, by = 10))) +
  scale_y_continuous(limits = c(0, 1000),breaks = seq(0, 1000, by = 200)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "B. Distribution of Age at First\n    Marriage",
       x = "Age at First Marriage",
       y = "")
# Plot for age at first birth
first_birth_plot <- gss_data %>%
  mutate(age_at_first_birth = floor(age_at_first_birth)) %>%
  select(age_at_first_birth) %>%
  ggplot(aes(x = age_at_first_birth)) +
  geom_bar(widh = 0.95, fill = "#4DAF4A") +
  scale_x_continuous(limits = c(15, 50), breaks = c(18, 45, seq(0, 50, by = 10))) +
  scale_y_continuous(limits = c(0, 1000),breaks = seq(0, 1000, by = 200)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "C. Distribution of Age at First Birth",
       x = "Age at First Birth",
       y = "")
# Plot years between marriage and birth
mar_and_birth <- gss_data %>% 
  mutate(age_at_first_marriage = floor(age_at_first_marriage), 
         age_at_first_birth = floor(age_at_first_birth)) %>% 
  mutate(years_btwn_mar_birth = age_at_first_birth - age_at_first_marriage) %>%
  group_by(years_btwn_mar_birth) %>% 
  summarise(n = n()) 
mar_birth_plot <- mar_and_birth %>%
  ggplot(aes(x = years_btwn_mar_birth, y = n)) + 
  geom_bar(stat = "identity", width = 0.95, fill = "#984EA3") +
  scale_x_continuous(limits = c(-29, 26),
                     breaks = seq(-30, 25, by = 5)) +
  scale_y_continuous(limits = c(0, 1100),breaks = seq(0, 1100, by = 100)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "A. Distribution of Years between\n    Marriage and Birth",
       x = "Age at First Birth - Age at First Marriage",
       y = "Number of Responses")
mar_birth_plot + (marriage_plot / first_birth_plot)
```

@@

```{r maplocation, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Immigrants Currently Residing in Canada by Province"}
immigrant_settled <- gss_data %>%
  filter(place_birth_canada == "Born outside Canada", place_birth_father == "Born outside Canada", place_birth_mother == "Born outside Canada") %>%
  group_by(province) %>%
  summarise(n = n())
# Code from https://cran.r-project.org/web/packages/mapcan/vignettes/choropleth_maps_vignette.html
pr_geographic <- mapcan(boundaries = province,
       type = standard)
pr_geographic <- full_join(pr_geographic, 
           immigrant_settled, 
           by = c("pr_english" = "province")) %>%
  mutate(n = replace_na(n, 0))
# graph labels
map_labels <- pr_geographic %>%
  group_by(pr_english) %>%
  summarise(mid_lat = mean(lat), mid_long = mean(long))
# fix position of certain labels #https://stackoverflow.com/questions/14782206/how-do-i-change-a-single-value-in-a-data-frame
map_labels[1, 3] = map_labels[1, 3] + 250000
map_labels[2, 3] = map_labels[2, 3] + 250000
map_labels[12, 3] = map_labels[12, 3] - 150000
map_labels[1, 2] = map_labels[1, 2] + 250000
map_labels[9, 2] = map_labels[9, 2] + 250000
map_labels[11, 3] = map_labels[11, 3] - 250000
pr_geographic <- inner_join(pr_geographic, map_labels, by = "pr_english")
# plot map
pr_geographic %>%
  ggplot(aes(x = long, y = lat, group = group, fill = n)) +
  geom_polygon() +
  coord_fixed() +
  theme_mapcan() +
  scale_fill_viridis_c(name = "Number of Immigrants") +
  geom_text(aes(label = pr_alpha, x = mid_long, y = mid_lat))
```

@@

# Plot for lang spoken
lang_spoke_data %>%
  ggplot(aes(x = reorder(province, n), fill = language_home)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(),
        legend.position = "bottom") +
  labs(x = "Province",
       y = "Number of responses",
       fill = "Language\nspoken in\nhousehold") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()
```

@@

```{r income, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Income"}

@@

```{r hoursworked, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Hours Worked"}

@@

# we can maybe look at hrs worked for a subset of the population i.e.
# location like "plot for place of birth"?
# Plot for hours worked
hrs_worked_data %>%
  ggplot(aes(x = average_hours_worked)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), 
        legend.position = "none") +
  labs(title = "Distribution of Average Number of Hours Worked per Week",
       x = "Average Number of Hours Worked per Week",
       y = "Number of Responses") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()

@@

Figure \@ref(fig:religion) shows the distribution of responses when respondents were asked to rate the importance of religion in their lives. Here we see that the religion is still important to the majority of Canadians, with the most common answer being "Very important", with almost 7000 responses, followed by "Somewhat important" with just over 6000 responses. Conversely, the least common response was "Don't Know", with around 200 responses. About 13,000 respondents answered that religion was either "Very important" or "Somewhat important", with around 7000 respondents answering that religion was either "Not very important" or "Not at all important". This means that while religion still plays an important role in the majority of Canadian's lives, it is of little importance to a large minority of Canadians.

```{r religion, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Religion"}

# Data wrangling for religion
religion_data <- gss_data %>%
  select(regilion_importance) %>% 
  add_count(regilion_importance) %>% 
  na.omit()

# we can maybe look at religion importance for a subset of the population i.e.
# location like "plot for place of birth"?
# Plot for religion
religion_data %>%
  ggplot(aes(x = reorder(regilion_importance, n))) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), 
        legend.position = "none") +
  labs(title = "Distribution of Religion Importance",
       x = "Religion Importance",
       y = "Number of Responses") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()

@@

@@

@@

```{r mentalhealth, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Mental Health"}

# Plot for mental health
mental_health_data %>%
  ggplot(aes(x = self_rated_mental_health, fill = age_grp)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank()) +
  labs(title = "Distribution of Self Rated Mental Health",
       x = "Self Rated Mental Health",
       y = "Number of Responses", 
       fill = "Age groups") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()

```
\newpage

@@

## Family Planning

Canada has always been committed to having strong family planning and sexual and reproductive health (SRH) services. In 2017, Global Affairs Canada -- a branch of the Canadian government -- dedicated CA$650 million over the next 3 years towards promoting the sexual and reproductive health and rights of women and girls [@cadleadership]. In 2015, Canada committed CA\$3.5 billion towards maternal, newborn and child health, including improving SRH services and information [@famplanning]. This strong foothold in the realm of family planning is reflected in this survey in a number of ways. For example, Figure \@ref(fig:householdsize) showcases the distribution of the size of Canadian households, broken down by what type of family lives in that household. Here we see that the majority of households only consist of 1 or 2 people, followed by households containing 3 or 4 people, of which there were a very similar amount. Notably, there is a steep drop-off in frequency of households larger than 4 people, with about a third as many 5-person households as there are 3 or 4-person, and less than a sixth as many households containing 6 or more people. This is likely a reflection of Canada's ample family planning resources leading couples more likely to opt for 1 or 2 children, thus keeping the average household size low. Figure \@ref(fig:numchildren) shows the distribution of the total number of children had by respondents. Interestingly, we see that by far the most frequent responses are 0 children and 2 children, of which there were a nearly identical number of responses. This means that once a couple has one child, they are then more likely to opt to have another child. This seemingly conflicts with Figure \@ref(fig:householdsize), which showed that 3-person households were just as common as 4-person households. There could be a couple reasons for these seemingly disjoint results. The firstly, relative to the general population, a high proportion of elderly people were surveyed. As elderly people are unlikely to still be living with their children, this would naturally lower the average household size without affecting the average number of children. As well, members of the same household are not necessarily parents or children, nor even family members. For example, a small proportion of 3-person households consist of couples. As these households are classified as couples and not intact families, this entails that the 3rd member of the household is not one of the couple's children. In Figure \@ref(fig:householdsize) we also see that aside from 0 and 2, 1 and 3 are both by far the next most common responses, with about 2900 respondents having 1 child and 3200 having 3. This also goes to show that Canada's continued investments into SRH and family planning services have been effective in keeping reproductive rates healthy and sustainable on both an individual and aggregate level.

@@

# Additional details

>>>>

---
title: "Say Cheese! A Snapshot of Canadian Family Demographics"
subtitle: "An Analysis of Insights from Canada's 2017 General Social Survey on Family"
author: 
  - Kimlin Chin
  - Jack McKay
  - Jacob Yoke Hong Si
thanks: "Code and data are available at: https://github.com/jmacattack27/General-Social-Survey-on-Family-2017- ."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "What does the average family look like in Canada? Using data from the 2017 Canadian General Social Survey (GSS) on Family, we look to paint a picture of Canadian families through the lens of family size and type, marriage and children, place of origin, income, religion and mental health. We found that most families are small, well-off and in good mental health, and that immigrants make up about one fifth of the population, with a propensity to settle in larger cities and urban centers. We also discuss possible factors that may have influenced this family image, including Canada's investment in family planning and immigration levels planning."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

```{r agesrespondents, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Distribution of Respondents' Ages"}

@@

# Plot for total number of children
five_year_groupings %>%
  ggplot(aes(x = age_group_5)) +
  geom_bar() +
  # scale_y_continuous(breaks = seq(0, 7000, by = 1000)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 50, hjust=1)) +
  labs(x = "Five-year Age groups",
       y = "Number of responses")

@@

In figure \@ref(fig:onepersonhouseholds), we take a closer look at the individuals living alone in terms of their ages and marital status. The shape of the distribution mirrors the shape of the age distribution of the entire sample, which implies that it was chosen randomly and appropriately. The percent of people living alone who are married is in the minority. We see the trend that the majority of persons less than 50 years are single, never married, and then this number remains constant until about 65 years when it starts to decrease. At 50 years and over there are more divorced persons, peaking at 65-69 years. The number of widowed individuals also increases steadily from 50 years and spikes at 80 years and over.

```{r onepersonhouseholds, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Age Distribution of people living alone"}

marital_status_data <- gss_data %>%
  select(marital_status, caseid)

d <- raw_data %>%
  mutate(household_size = case_when(hsdsizec == 6 ~ "6+",
                                    TRUE ~ as.character(hsdsizec)),
         age_group_5 = case_when(agegr5 == 1 ~ "15 to 19 years",
                                 agegr5 == 2 ~ "20 to 24 years",
                                 agegr5 == 3 ~ "25 to 29 years",
                                 agegr5 == 4 ~ "30 to 34 years",
                                 agegr5 == 5 ~ "35 to 39 years",
                                 agegr5 == 6 ~ "40 to 44 years",
                                 agegr5 == 7 ~ "45 to 49 years",
                                 agegr5 == 8 ~ "50 to 54 years",
                                 agegr5 == 9 ~ "55 to 59 years",
                                 agegr5 == 10 ~ "60 to 64 years",
                                 agegr5 == 11 ~ "65 to 69 years",
                                 agegr5 == 12 ~ "70 to 74 years",
                                 agegr5 == 13 ~ "75 to 79 years",
                                 agegr5 == 14 ~ "80 years and over"))

d <- left_join(d, marital_status_data, by = c("pumfid" = "caseid")) %>%
  na.omit()

d %>%
  filter(household_size == 1) %>%
  ggplot(aes(x = age_group_5, fill = marital_status)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(x = "Age of respondent",
       y = "Number of responses",
       fill = "Marital Status") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 50, hjust=1))

@@

Figure \@ref(fig:numchildren) shows the distribution of the total number of children per respondent and their marital status. Note that this is not *per household* i.e. persons living together, but rather is the number of children had by the respondent of the household throughout their lifetime. Here we see that the majority of respondents either had 0 children or 2, with almost an identical number of respondents with 0 children as there are with 2. Similarly, the number of respondents with 1 child is very close to the number of those with 3. The majority of married couples have children, and most people who have zero children are single, never married. We note that about `r round(zero_children$total_percent,2)`% of Canadians do not have any children, and those who do usually have 1-3 children.

```{r numchildren, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Total Number of Children per Respondent with Marital Status"}
# Plot for total number of children
child_data %>%
  ggplot(aes(x = total_children1, fill = marital_status)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 7000, by = 1000)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Total Number of Children per Respondent",
       x = "Number of Children",
       y = "Number of Responses",
       fill = "Marital Status") +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r agemarbirth, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Ages at Marriage and Birth"}
# Plot for age at first marriage
marriage_plot<- gss_data %>%
  mutate(age_at_first_marriage = floor(age_at_first_marriage)) %>%
  select(age_at_first_marriage) %>%
  ggplot(aes(x = age_at_first_marriage)) +
  geom_bar(width = 0.95, fill = "#E41A1C") +
  scale_x_continuous(limits = c(15, 50), breaks = c(15, seq(0, 50, by = 10))) +
  scale_y_continuous(limits = c(0, 1000),breaks = seq(0, 1000, by = 200)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "B. Distribution of Age at First\n    Marriage",
       x = "Age at First Marriage",
       y = "")
# Plot for age at first birth
first_birth_plot <- gss_data %>%
  mutate(age_at_first_birth = floor(age_at_first_birth)) %>%
  select(age_at_first_birth) %>%
  ggplot(aes(x = age_at_first_birth)) +
  geom_bar(widh = 0.95, fill = "#4DAF4A") +
  scale_x_continuous(limits = c(15, 50), breaks = c(18, 45, seq(0, 50, by = 10))) +
  scale_y_continuous(limits = c(0, 1000),breaks = seq(0, 1000, by = 200)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "C. Distribution of Age at First\nBirth",
       x = "Age at First Birth",
       y = "")
# Plot years between marriage and birth
mar_and_birth <- gss_data %>% 
  mutate(age_at_first_marriage = floor(age_at_first_marriage), 
         age_at_first_birth = floor(age_at_first_birth)) %>% 
  mutate(years_btwn_mar_birth = age_at_first_birth - age_at_first_marriage) %>%
  group_by(years_btwn_mar_birth) %>% 
  summarise(n = n()) 
mar_birth_plot <- mar_and_birth %>%
  ggplot(aes(x = years_btwn_mar_birth, y = n)) + 
  geom_bar(stat = "identity", width = 0.95, fill = "#984EA3") +
  scale_x_continuous(limits = c(-29, 26),
                     breaks = seq(-30, 25, by = 5)) +
  scale_y_continuous(limits = c(0, 1100),breaks = seq(0, 1100, by = 100)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "A. Distribution of Years between\n    Marriage and Birth",
       x = "Age at First Birth - Age at First Marriage",
       y = "Number of Responses")
mar_birth_plot + (marriage_plot / first_birth_plot)
```

@@

```{r maplocation, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Immigrants Currently Residing in Canada by Province"}
immigrant_settled <- gss_data %>%
  filter(place_birth_canada == "Born outside Canada", place_birth_father == "Born outside Canada", place_birth_mother == "Born outside Canada") %>%
  group_by(province) %>%
  summarise(n = n())
# Code from https://cran.r-project.org/web/packages/mapcan/vignettes/choropleth_maps_vignette.html
pr_geographic <- mapcan(boundaries = province,
       type = standard)
pr_geographic <- full_join(pr_geographic, 
           immigrant_settled, 
           by = c("pr_english" = "province"))
# graph labels
map_labels <- pr_geographic %>%
  group_by(pr_english) %>%
  summarise(mid_lat = mean(lat), mid_long = mean(long))
# fix position of certain labels #https://stackoverflow.com/questions/14782206/how-do-i-change-a-single-value-in-a-data-frame
map_labels[1, 3] = map_labels[1, 3] + 250000
map_labels[2, 3] = map_labels[2, 3] + 250000
map_labels[12, 3] = map_labels[12, 3] - 150000
map_labels[1, 2] = map_labels[1, 2] + 250000
map_labels[9, 2] = map_labels[9, 2] + 250000
map_labels[11, 3] = map_labels[11, 3] - 250000

pr_geographic <- inner_join(pr_geographic, map_labels, by = "pr_english") %>%
  mutate(color = case_when(pr_alpha %in% c("SK", "NL") ~ "#ffffff",
                           TRUE ~ "black"))
nb_position <- pr_geographic %>%
  filter(pr_alpha == "NB") %>%
  select(mid_long, mid_lat)

nb_position <- c(unique(nb_position$mid_long), unique(nb_position$mid_lat))

ns_position <- pr_geographic %>%
  filter(pr_alpha == "NS") %>%
  select(mid_long, mid_lat)

ns_position <- c(unique(ns_position$mid_long), unique(ns_position$mid_lat))

pe_position <- pr_geographic %>%
  filter(pr_alpha == "PE") %>%
  select(mid_long, mid_lat)

pe_position <- c(unique(pe_position$mid_long), unique(pe_position$mid_lat))

pr_geographic <- pr_geographic %>%
  mutate(pr_alpha = case_when(pr_alpha %in% c("ON", "AB", "SK", "NL", "BC", "MB", "QC", "NU", "YT", "NT") ~ pr_alpha))

# plot map
pr_geographic %>%
  ggplot(aes(x = long, y = lat, group = group, fill = n)) +
  geom_polygon() +
  coord_fixed() +
  theme_mapcan() +
  scale_fill_viridis_c(name = "Number of Immigrants") +
  scale_color_manual(values = c("#ffffff","#000000", "#000000")) +
  geom_text(aes(label = pr_alpha, x = mid_long, y = mid_lat, color = color)) +
  geom_text(aes(label = "NB", x = nb_position[1], y = nb_position[2]), nudge_y = -250000, nudge_x = -100000) +
  geom_text(aes(label = "NS", x = ns_position[1], y = ns_position[2]), nudge_x = 250000) +
  geom_text(aes(label = "PE", x = pe_position[1], y = pe_position[2]), nudge_y = 150000) +
  guides(color=FALSE) # https://stackoverflow.com/questions/42686475/how-do-i-just-remove-one-of-the-two-legends-in-ggplot
```

@@

# Plot for lang spoken
lang_spoke_data %>%
  ggplot(aes(x = reorder(province, n), fill = language_home)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(),
        legend.position = "bottom") +
  guides(fill=guide_legend(nrow=4, byrow=TRUE)) +
  labs(x = "Province",
       y = "Number of responses",
       fill = "Language\nspoken in\nhousehold") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()
```

@@

```{r income, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Family Income with Household Size"}

@@

```{r hoursworked, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Average Hours Worked per Week"}

@@

hrs_worked_summary <- hrs_worked_data %>%
  group_by(average_hours_worked) %>%
  summarise(percent = 100*n()/nrow(hrs_worked_data))

# 5+ people in a houshold percentage
hrs_freq <- hrs_worked_summary[3,] %>%
  select(percent) %>%
  summarise(total_percent = sum(percent))

# we can maybe look at hrs worked for a subset of the population i.e.
# location like "plot for place of birth"?
# Plot for hours worked
hrs_worked_data %>%
  ggplot(aes(x = average_hours_worked)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar(fill = "#4DAF4A") +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), 
        legend.position = "none") +
  labs(title = "Distribution of Average Number of Hours Worked per Week",
       x = "Average Number of Hours Worked per Week",
       y = "Number of Responses") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()

@@

```{r religion, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Religion affiliation and importance"}

# Data wrangling for religion
religion_data <- raw_data %>%
  select(religflg, pumfid, agegr5) %>%
  mutate(religious_affiliation = case_when(religflg == 1 ~ "Has a religious affiliation",
                                           religflg == 2 ~ "No religious affiliation",
                                           religflg == 7 ~ "Don't know",
                                           religflg == 8 ~ "Refusal",
                                           religflg == 9 ~ "Not stated"),
         age_group_5 = case_when(agegr5 == 1 ~ "15 to 19 years",
                                 agegr5 == 2 ~ "20 to 24 years",
                                 agegr5 == 3 ~ "25 to 29 years",
                                 agegr5 == 4 ~ "30 to 34 years",
                                 agegr5 == 5 ~ "35 to 39 years",
                                 agegr5 == 6 ~ "40 to 44 years",
                                 agegr5 == 7 ~ "45 to 49 years",
                                 agegr5 == 8 ~ "50 to 54 years",
                                 agegr5 == 9 ~ "55 to 59 years",
                                 agegr5 == 10 ~ "60 to 64 years",
                                 agegr5 == 11 ~ "65 to 69 years",
                                 agegr5 == 12 ~ "70 to 74 years",
                                 agegr5 == 13 ~ "75 to 79 years",
                                 agegr5 == 14 ~ "80 years and over")) %>%
  na.omit()

religion_data <- full_join(religion_data, gss_data, by = c("pumfid" = "caseid"))

# we can maybe look at religion importance for a subset of the population i.e.
# location like "plot for place of birth"?
# Plot for religion
religion_data %>%
  ggplot(aes(x = religious_affiliation, fill = regilion_importance)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank()) +
  labs(x = "Religious Affiliation",
       y = "Number of Responses",
       fill = "Religion Importance") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()

@@

Figure \@ref(fig:religion) shows if respondents have a religious affiliation and how important they think that religion is. The majority of respondents have a religious affiliation. For most respondents who have a religion, religion is at least somewhat important to them, although there is a non-negligible proportion of these persons who do not think that religion is important. Similarly, for persons with no religious affiliation, the majority think that religion is not very important at all, but there are still some who believe in the importance of religion. This means that while religion still plays an important role in the majority of Canadian's lives, it is of little importance to a large minority of Canadians.

```{r religionages, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Religion importance by age group"}

religion_data %>%
  mutate(religion_important = case_when(regilion_importance %in% c("Somewhat important", "Very important") ~ "Somewhat - Very important",
                                        regilion_importance %in% c("Not very important", "Not at all important") ~ "Not very - Not at all important",
                                        TRUE ~ regilion_importance)) %>%
  ggplot(aes(x = age_group_5, fill = religion_important)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(x = "Age of respondent",
       y = "Number of responses",
       fill = "Religion importance") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 50, hjust=1))

Figure \@ref(fig:religionages) shows religion importance by age group. We combined categories of religion importance from the previous graph into "Somewhat - Very important", "Not very - Not at all important" and "Don't Know" to get a clearer picture of important vs not important. It can be seen that as age increases, the proportion of persons who think religion is important increases. Notably for individuals less than 30 years old, an equal if not larger percent think that religion is not very important or not important at all.

@@

```{r mentalhealth, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of Self-rated Mental Health for diffrent age groups"}

# Plot for mental health
mental_health_data %>%
  ggplot(aes(x = self_rated_mental_health, fill = age_grp)) + # https://datavizpyr.com/re-ordering-bars-in-barplot-in-r/
  geom_bar() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank()) +
  labs(x = "Self Rated Mental Health",
       y = "Number of Responses", 
       fill = "Age groups") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip()

@@

@@

In this paper, we explored different aspects of what makes a Canadian family - number of people, their relationships to each other through marriage and children, which lands they have immigrated from, how long they work and how much money they make, their religious beliefs and how they feel about their mental health. We now discuss more in-depth each of these characteristics and how different powers may have influenced how so it came to be.

## Family Sizes and Family Planning

Canada has always been committed to having strong family planning and sexual and reproductive health (SRH) services. In 2017, Global Affairs Canada -- a branch of the Canadian government -- dedicated CA$650 million over the next 3 years towards promoting the sexual and reproductive health and rights of women and girls [@cadleadership]. In 2015, Canada committed CA\$3.5 billion towards maternal, newborn and child health, including improving SRH services and information [@famplanning]. This strong foothold in the realm of family planning is reflected in this survey in a number of ways. For example, Figure \@ref(fig:householdsize) showcases the distribution of the size of Canadian households, broken down by what type of family lives in that household. Here we see that the majority of households only consist of 1 or 2 people, followed by households containing 3 or 4 people, of which there were a very similar amount. Notably, there is a steep drop-off in frequency of households larger than 4 people, with about a third as many 5-person households as there are 3 or 4-person, and less than a sixth as many households containing 6 or more people. This is likely a reflection of Canada's ample family planning resources leading couples more likely to opt for 1 or 2 children, thus keeping the average household size low.

Figure \@ref(fig:numchildren) shows the distribution of the total number of children had by respondents. Interestingly, we see that by far the most frequent responses are 0 children and 2 children, of which there were a nearly identical number of responses. This means that once a couple has one child, they are then more likely to opt to have another child. This seemingly conflicts with Figure \@ref(fig:householdsize), which showed that 3-person households were just as common as 4-person households. However, the difference is because the total number of children in figure \@ref(fig:numchildren) is not necessarily the number of children living in the household. More than half of the respondents were 50 years and over, so they are unlikely to be living with their children. As well, members of the same household are not necessarily parents or children, nor even family members. For example, a small proportion of 3-person households consist of couples. As these households are classified as couples and not intact families, this entails that the 3rd member of the household is not one of the couple's children. In Figure \@ref(fig:householdsize) we also see that aside from 0 and 2, 1 and 3 are both by far the next most common responses, with about 2900 respondents having 1 child and 3200 having 3. This also goes to show that Canada's continued investments into SRH and family planning services have been effective in keeping reproductive rates healthy and sustainable on both an individual and aggregate level.

@@

The supplementary survey can be found at https://forms.gle/jun4vZaAh247442s5 .
